%!TEX root = /Users/nunolourenco/Documents/FCTUC/Mestrado/2010_2011/Thesis/Thesis/thesis.tex
\chapter{Implementation Details}
\label{chap:implementation}

Since this dissertation is part of the Masters Degree in Informatics Engineering, it is important to refer some technical aspects about the implementation. In this chapter we give a brief description of the technical choices made, and we present a quick overview of the entire system.

\section{Technical Choices}

To implement the first versions of the system we decide to use the Python language. One of the reasons for this is related with the simplicity of the language, and easiness in making changes. The other reason is related with the L-BFGS method, which is written in Fortran. With python it is easy to connect modules made in Fortran with our code.

After we have a stable version, we decided to write it in C. The reason for this was the performance of the python language. 

The C language, despite being very fast, has some programming overhead, when compared, for instance, with python. During the conversion of the python code to C, some bugs appeared, as one should expect. They were solved using the \emph{gdb} debugging system.

We used \emph{Git} as revision control system. As online repository our choice was the \emph{Github} platform, with private repositories.

\section{Architecture Overview}
We tried to keep our architecture simple and modular. In Fig.\ref{fig:class_diagram} we present a class diagram, which in our opinion gives a better insight in how the system is organized.

\botapic[0.7]{class_diagram}{CGACO Class Diagram}
\textbf{DACCO} – Main class that runs the ACO algorithm for a certain number of populations.

\textbf{Population} – Class that contains a certain number of individuals, with common properties.

\textbf{Ant} – Class that simulates the real ant. It is responsible for the individual methods of each ant. The main methods are:
\begin{itemize}
	\item Construction of individual solutions:
		\begin{itemize}
			\item Choose feasible neighborhoods;
			\item Apply a probabilistic rule to choose which components should be added to the solutions.
		\end{itemize}
	\item It calls the Fitness Function, in order to assess the quality of the solution that has been built;
	\item Mapping of the continuous solutions to the discrete space;
	\item Discrete local search. This local search works in the search space, and tries to improve the ant solutions applying random changes to the current solutions.
\end{itemize}

\textbf{Utils} – Class that has a set of procedures to help all the other classes accomplish their individual tasks. Examples of these methods are: the euclidean distance method, that receives to points in the space and determines their euclidean distance; the method to read the parameters file; methods to write data to files, in order to allow us to see what is happening in the optimization process.

\textbf{Fitness Function} – Class that is responsible for evaluate the quality of the individuals that belong to our system. This class receives an ant solution, that is already in the continuous space, and it performs the evaluation of the solutions. The evaluation is performed by a Fortran function. In order to use it, we applied two different techniques:
	\begin{itemize}
		\item To connect the Fortran code with C we created several object files and the we compiled them together. One aspect that we had to take into account was that sometimes the Fortran functions have and ``\_" after the name.
		\item To connect Fortran with Python, we created a static library, using the Cython \footnote{http://cython.org/} framework.
		\end{itemize}

